---
title       : "Shiny App"
subtitle    : "Tokyo Population Map"
author      : "Fritz Lin"
framework   : revealjs        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

<style>
code {
  color: red;
}

strong {
  font-weight: 900;
  color: #42f48f;
}
</style>

## Developing Data Products
---------------------

### Shiny App: Tokyo Population Map

Coursera Data Science Specialization  
Fritz Lin  
September, 21st 2017

---  

## Introduction

* This presentation is part of the [Coursera course on Developing Data Products](https://www.coursera.org/learn/data-products) 

* The **Shiny app** pitched by this presentation is at
https://alpenfritz.shinyapps.io/tokyopopmap/

* The Shiny app **source code** is available at
https://github.com/alpenfritz/tokyopopmap/tree/master/app

--- 

## Tokyo Population Map App

* The [Tokyo Population Map App](https://alpenfritz.shinyapps.io/tokyopopmap/) shows the population size in each of the 23 wards of Tokyo
    - population size of each ward is further subdivided by *gender* and *age*

* **Interactive Map** using `leaflet` package to show information
    - population sizes of wards highlighted by a **choeroplath map**
    - using **popups** to display information on each ward 

* **Bar graphs** for comparing general population sizes or population sizes by gender and age
    - possible to filter depending on population size using a **slider**

--- 

## Interactive Map
```{r, echo = FALSE}
# Load static data
tWards <- readRDS('app/data/twards84.Rds')
```


```{r, echo = FALSE, warning=FALSE, message=FALSE, results='asis'}
library(leaflet)
library(rgdal)
library(htmlwidgets)
library(knitr)

# Find the edges of our map
bounds<-bbox(tWards)

pal <- colorBin("BuPu", tWards$POP_2015)

myPopup <- paste0("<strong>",tWards$CITY_ENG,"</strong> ",
                    tWards$SIKUCHOSON,
                    "<br><strong>Population:</strong> ",
                    formatC(tWards$POP_2015, format="d", big.mark = ","),
                    "<br><strong>Male:</strong> ",
                    formatC(tWards$M_2015, format="d", big.mark = ","),
                    "<br><strong>Female:</strong> ",
                    formatC(tWards$F_2015, format="d", big.mark = ","),
                    "<br><strong>15未満:</strong> ",
                    formatC(tWards$`15未満_2015`, format="d", big.mark = ","),
                    "<br><strong>15-64:</strong> ",
                    formatC(tWards$`15-64_2015`, format="d", big.mark = ","),
                    "<br><strong>65以上:</strong> ",
                    formatC(tWards$`65以上_2015`, format="d", big.mark = ","))

l <- leaflet() %>%
      addProviderTiles("Esri.WorldStreetMap", group = "Esri.WorldStreetMap") %>%
      addProviderTiles("Esri.WorldImagery", group = "Esri.WorldImagery") %>%
      addProviderTiles("OpenStreetMap.Mapnik", group = "OpenStreetMap.Mapnik") %>%
      addProviderTiles("OpenStreetMap.BlackAndWhite", group = "OpenStreetMap.BlackAndWhite") %>%
      # Centre the map in the middle of our co-ordinates
      setView(mean(bounds[1,]),
              mean(bounds[2,]),
              zoom=11
      ) %>%
      addPolygons(data = tWards,
                  color = "#000066", opacity = 1, weight = 2, group = "Choropleth Map",
                  fillColor = ~pal(POP_2015), fillOpacity = 0.7,
                  label=~paste0(CITY_ENG, " ", SIKUCHOSON),
                  labelOptions= labelOptions(direction = 'auto', textsize="15px",
                                             padding = "3px 8px"),
                  layerId=~JCODE,
                  popup = myPopup,
                  highlightOptions = highlightOptions(color = "#666", weight = 5,
                                                      bringToFront = TRUE)) %>%
      addLegend(title = "Population Size", pal = pal, values=tWards@data$POP_2015,
                opacity = 0.6, position="bottomright", group = "Choropleth Map") %>%
      addLayersControl(baseGroups = c("Esri.WorldStreetMap",
                                      "Esri.WorldImagery",
                                      "OpenStreetMap.Mapnik",
                                      "OpenStreetMap.BlackAndWhite"),
                       overlayGroups = c("Choropleth Map"),
                       options = layersControlOptions(collapsed = TRUE, autoZIndex = FALSE))

# Move to img folder
setwd("./assets/img")
saveWidget(l, "leaflet.html") # Save widget html
setwd("../..")

# Source saved file
cat('<iframe src="./assets/img/leaflet.html" width=100% height=500px allowtransparency="true"> </iframe>')
```

--- 

## Bar Graphs

```{r bargraphs, echo = FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=13, fig.align="center"}
library(ggplot2)
library(gridExtra)
library(dplyr)
library(tidyr)

g1 <- ggplot(tWards@data, aes(reorder(CITY_ENG,-POP_2015), POP_2015/1e3)) + 
        geom_bar(stat="identity", fill="steelblue", width = 0.6) + 
        coord_flip() +
        labs(x = "", y = "")

# subsetting
df <- subset(tWards@data, POP_2015>=4e5 & POP_2015<=1e6)
df <- dplyr::select(df, matches("JCODE|CITY_ENG|^M_2015|^F_2015"))
df <- gather(df, GENDER, POP, -JCODE, -CITY_ENG)
df$GENDER <- gsub(pattern = "_2015", replacement = "", df$GENDER)

g2 <- ggplot(df, aes(reorder(CITY_ENG,-POP), POP/1e3, fill=GENDER)) + 
          scale_fill_manual(values = c("#c43a3d", "steelblue")) +
          geom_bar(stat="identity", position = position_dodge(), width = 0.8) + 
          coord_flip() +
          labs(x = "", y = "") +
          theme(legend.position = c(0.8, 0.9))

# subsetting
df <- subset(tWards@data, POP_2015>=4e5 & POP_2015<=6e5)
df <- dplyr::select(df, matches("JCODE|CITY_ENG|^15未満_2015|^15-64_2015|^65以上_2015"))
df <- gather(df, AGE, POP, -JCODE, -CITY_ENG)
df$AGE <- gsub(pattern = "_2015", replacement = "", df$AGE)
      
g3 <- ggplot(df, aes(reorder(CITY_ENG,-POP), POP/1e3, fill=AGE)) +
          scale_fill_manual(values = c("steelblue", "#c43a3d", "#47892e")) +
          geom_bar(stat="identity", position = position_dodge(), width = 0.8) + 
          coord_flip() +
          labs(x = "", y = "") +
          theme(legend.position = c(0.8, 0.9))

grid.arrange(g1,g2,g3, ncol=3)
```
